---
title: "Korea_Kyushu_analysis"
author: "OEC"
date: "27/02/2024"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r}
#define theme_pub to apply to all 
setwd("/Users/olcraig/Desktop/KorKyu_project/R_script")
theme_Pub <- function(base_size=10, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(),
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "right",
               legend.direction = "vertical",
               legend.key.size= unit(0.5, "cm"),
               legend.spacing = unit(0, "cm"),
               legend.title = element_text(NULL),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))

}
```


```{r}
# @title
# Read data from INTERNET
library('googlesheets4')
 #gs4_deauth() # if needed
 gs4_auth(email='oliver.craig@york.ac.uk') # only if needed
# gs4_auth(email='jasmine.lundy@york.ac.uk') # only if needed
Data <-read_sheet('https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0', sheet="Data") |> as.data.frame()
library(dplyr)
Data

```


#1. Preservation and summary statisitcs
#1a. Preservation by site

```{r}
library(dplyr)
#Grouped by site
Conc_site <- Data %>% filter(Type == 'Ceramic') %>%
  group_by(Site_code) %>% 
  summarize(n = n(), mean.conc = mean(Concentration),
            sd.conc = sd(Concentration), n_lipid = sum(Concentration >=5))%>%
  mutate("%interp"= n_lipid/n*100)
Conc_site <- Conc_site %>% arrange(factor(Site_code, levels = c('OBD', 'RMM', 'SHC', 'SAS','NBK', 'MSR', 'DPR', 'SDR', 'SWR', 'JKR', 'KCR', 'MJR')))
Conc_site
write.csv (Conc_site,"Conc.csv")

#Grouped by country
#Data %>% 
 # group_by(Country) %>% 
  #summarize(median.conc = median(Concentration),
           # mean.conc = mean(Concentration),
           # sd.conc = sd(Concentration),
            #n.conc = n())%>%
  #group_modify(~as.data.frame(t(quantile(.$Concentration))))

  #mutate(se.conc = sd.conc / sqrt(n.conc),
         #lower.ci.conc = mean.conc - qt(1 - (0.05 / 2), n.conc - 1) * se.conc,
         #upper.ci.conc = mean.conc + qt(1 - (0.05 / 2), n.conc - 1) * se.conc)
```
  
#1b. Preservation by Country

```{r}
Region <-Data %>%filter(Type == 'Ceramic') %>% 
 group_by(Country) %>% 
  summarize(n = n(), median.conc = median(Concentration),
           mean.conc = mean(Concentration),
           sd.conc = sd(Concentration), n_lipid = sum(Concentration >=5)) %>%
  mutate("%interp"= n_lipid/n*100)
Region
 write.csv (Region,"Region_preservation.csv")
#%>%
 # group_modify(~as.data.frame(t(quantile(.$Concentration))))%>%
  #mutate(se.conc = sd.conc / sqrt(n),
   #      lower.ci.conc = mean.conc - qt(1 - (0.05 / 2), n.conc - 1) * se.conc,
    #     upper.ci.conc = mean.conc + qt(1 - (0.05 / 2), n.conc - 1) * se.conc)

```
  
#2a.Millet identification by site by interpretable lipid

```{r}
Inter_lipid <- Data %>%filter (Concentration >=5)
Biomarker <-Inter_lipid %>% group_by(Site_code) %>% 
  summarize (n = n(), n_millet = sum (Millet == "yes"), n_aquatic = sum (Aquatic == "yes")) %>%
  mutate("%millet"= n_millet/n*100, "%aqatiuc"= n_aquatic/n*100)
Biomarker <- Biomarker %>% arrange(factor(Site_code, levels = c('OBD', 'RMM', 'SHC', 'SAS', 'MSR', 'DPR', 'SDR', 'SWR', 'JKR', 'KCR', 'MJR')))
Biomarker
write.csv (Biomarker,"Biomarker_preservation.csv")
```


#2b.Aquatic identification by site by interpretable lipid
```{r}

Millet <- Biomarker %>% arrange(factor(Site_code, levels = c('OBD', 'RMM', 'SHC', 'SAS', 'NBK', 'MSR', 'DPR', 'SDR', 'SWR', 'JKR', 'KCR', 'MJR')))
Millet
write.csv (Millet,"Marine_preservation.csv")
```



#3a. Isotope data overview and analysis


```{r}
library("ggplot2")
Iso_data <- Data %>% filter(d13C16 < 0)
p <-ggplot()+
  labs(y=expression(delta^{13}*C[18:0]*"(\u2030)"), x=expression(delta^{13}*C[16:0]*"(\u2030)"))+
  #scale_x_continuous(limits=c(-35,-10))+
  #scale_y_continuous(limits=c(-35,-10))+
  geom_point(data = Iso_data, aes(y=d13C18, x=d13C16, fill=Phase_2, size=Concentration), colour="black", pch=21)+
  coord_fixed(ratio = 1)+
  theme_Pub()
p  +   facet_wrap(~Country)
```

#3b.Isotpic Reference data

Obtain data for isotope references and concentations
```{r}
#Otain data for reference fats
library(dplyr)
Iso_ref <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0", sheet = "Iso_ref")|> as.data.frame()
Conc_ref <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0", sheet = "Conc_ref")|> as.data.frame()
```

Find means and ranges of isotope and concentrations values
```{r}
Iso_sum <- Iso_ref %>% group_by(Category) %>% summarize (Meand13C16 = mean(d13C16), SDd13C16 = sd(d13C16), Meand13C18 = mean(d13C18), SDd13C18 = sd(d13C18), n = n())
Conc_sum <- Conc_ref %>% group_by(Category) %>% summarize (Concd13C16 = mean(Ab_Percent_C16), SDC16 = sd(Ab_Percent_C16), Concd13C18 = mean(Ab_Percent_C18), SDC18 = sd(Ab_Percent_C18), n_conc = n())
Iso_sum
Conc_sum
```

#3c Generate arandom concentration values for each isotope value based on the mean and sd of the concentration 

```{r}
library(dplyr)
library(purrr)
library(tidyr)
Conc <- inner_join(Iso_sum, Conc_sum, 
           by = c("Category"))
Conc
# Generate the random values using purrr::map and bind them into a single column
Conc <- Conc %>%
  # Use purrr::pmap to generate random numbers based on each row's parameters
  mutate(
    Concd13C16 = pmap(list(Concd13C16, SDC16, n), function(mean, sd, n) {
      rnorm(n, mean = mean, sd = sd)
    }),
    Concd13C18 = pmap(list(Concd13C18, SDC18, n), function(mean, sd, n) {
      rnorm(n, mean = mean, sd = sd)
    })
  ) %>%
  unnest(cols =c(Concd13C16, Concd13C18))
  # Unnest the list column to create one long column of generated values
```

#4. Create a mixing model with MixSiar
#4a Load the consumer data
```{r}
#remotes::install_github("brianstock/MixSIAR")
library(MixSIAR)

Pots <- Data %>% filter(d13C16 < 0) %>% dplyr::select(Sample_ID, Site_name, Region, Phase_2, d13C16, d13C18)
write.csv (Pots,"/Users/olcraig/Documents/GitHub/KoreavsKyushu/pots.csv")
mix <- load_mix_data (filename="/Users/olcraig/Documents/GitHub/KoreavsKyushu/pots.csv", 
                     iso_names=c("d13C16", "d13C18"), 
                     factors="Sample_ID",
                     fac_random=FALSE, 
                     fac_nested=FALSE, 
                     cont_effects=NULL)
mix
```


```{r}
#Install the latest version of MixSIAR from github!
#remotes::install_github("brianstock/MixSIAR",  auth_token = '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', force = TRUE,  dependencies=T) 
```

#4b. Create a source data file

```{r}
#Select categories for model and index
library(MixSIAR)
Iso_ref1 <- Iso_ref %>% filter(Category == "Rice" | Category == "Millet" | Category == "Wild ruminant" | Category == "Marine") %>% mutate ("Sources"= Category) %>%dplyr::select(Sources, d13C16, d13C18) 
Iso_ref1 <- Iso_ref1 %>%
  arrange(Sources) %>%
  mutate(Observaton = row_number()) %>%
  ungroup()
Iso_ref1
#Repeate for concentration
Conc_ref1 <- Conc %>% filter(Category == "Rice" | Category == "Millet" | Category == "Wild ruminant" | Category == "Marine") %>% mutate ("Sources"= Category) %>%dplyr::select(Sources, Concd13C16, Concd13C18) 
Conc_ref1 <- Conc_ref1 %>%
  arrange(Sources) %>%
  mutate(Observaton = row_number()) %>%
  ungroup()
Conc_ref1
#Join. the data togeter
Source <- Source <- inner_join(Conc_ref1, Iso_ref1, by = c("Observaton"))
Source
```

Load the source data

```{r}
write.csv (Source,"/Users/olcraig/Documents/GitHub/KoreavsKyushu/source.csv", row.names=FALSE)
source <- load_source_data (filename="/Users/olcraig/Documents/GitHub/KoreavsKyushu/source.csv", 
                           source_factors=NULL,
                             conc_dep=TRUE, 
                           data_type="raw", 
                           mix)
```
#4c Load the TDFs if relevant or neill return
```{r}
# Load the discrimination/TDF data
discr <- load_discr_data(filename="/Users/olcraig/Documents/GitHub/KoreavsKyushu/scr.csv", mix)
```
```{r, eval=FALSE}
# Make an isospace plot
plot <- plot_data(filename="isospace_plot", return_obj=T, mix,source,discr)
```
```{r}
# Calculate the convex hull area, standardized by source variance
calc_area(source=source,mix=mix,discr=discr)
```

```{r, eval=FALSE}
# default "UNINFORMATIVE" / GENERALIST prior (alpha = 1)
#plot_prior(alpha.prior=1,source)
```

#4d. Create the JAGS file and run the model
```{r, eval=FALSE}
# Write the JAGS model file
model_filename <- "MixSIAR_model.txt"   # Name of the JAGS model file
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix, source)
```
```{r, eval=FALSE}
#run <- list(chainLength=200000, burn=150000, thin=50, chains=3, calcDIC=TRUE)
```
```{r, eval=FALSE}
jags.1 <- run_model(run="test", mix, source, discr, model_filename)
```
```{r, eval=FALSE}
jags.1 <- run_model(run="normal", mix, source, discr, model_filename)
```

#4d. Obtain the model outputs
```{r, eval=FALSE}
output_options <- list(summary_save = TRUE,
                       summary_name = "summary_final",
                       sup_post = TRUE,
                       plot_post_save_pdf = TRUE,
                       plot_post_name = "posterior_density",
                       sup_pairs = TRUE,
                       plot_pairs_save_pdf = TRUE,
                       plot_pairs_name = "pairs_plot",
                       sup_xy = TRUE,
                       plot_xy_save_pdf = TRUE,
                       plot_xy_name = "xy_plot",
                       gelman = TRUE,
                       heidel = TRUE,
                       geweke = TRUE,
                       diag_save = TRUE,
                       diag_name = "diagnostics",
                       indiv_effect = FALSE,
                       plot_post_save_png = TRUE,
                       plot_pairs_save_png = TRUE,
                       plot_xy_save_png = TRUE,
                       diag_save_ggmcmc = FALSE,
                       return_obj=TRUE)
```


```{r}
# Get posterior chains into one tidy data frame
library(R2jags)

summary(jags.1)
attach.jags(jags.1)
```
Run diagnostics and create a data frame with some summary stats. 

```{r}
diag <- output_diagnostics(jags.1, mix, source, output_options)
head(diag$gelman)
df.stats <- output_stats(jags.1, mix, source, output_options)
rownames(df.stats)
df.stats <- as.data.frame(df.stats)

```


```{r}
df.stats <- cbind(Sample_category = rownames(df.stats), df.stats)
rownames(df.stats) <- 1:nrow(df.stats)
df.stats
```


```{r}
library(tidyr)
df.stats <- df.stats %>% separate_wider_delim (Sample_category, ".", names = c("number", "Sample_ID", "Category"))
colnames(df.stats)[11] <- "p95"
colnames(df.stats)[7] <- "p5"
df.stats
```


Join with isotope data
```{r}

Data_pots <- inner_join(Iso_data, df.stats, 
           by = c("Sample_ID"))
Data_pots
```

Output posteriors 

```{r}
g.post <- output_posteriors(jags.1, mix, source, output_options)
```


#5. Create biomarker plots 

Create a base plot
```{r}

p <-ggplot()+
  labs(y=expression(delta^{13}*C[18:0]*"(\u2030)"), x=expression(delta^{13}*C[16:0]*"(\u2030)"))+
  scale_x_continuous(limits=c(-37,-15))+
  scale_y_continuous(limits=c(-37,-15))+
     stat_ellipse(geom= "polygon", data=Iso_ref1, alpha=0.2, colour="lightgrey",size=0.1,
               aes (x = d13C16, y = d13C18, group = Sources),
              level = 0.68)+
annotate("text", label = "C4/millet", x = -20, y = -17, size = 3, colour = "black")+ 
  annotate("text", label = "Marine", x = -25, y = -22, size = 3, colour = "black")+ 
  #annotate("text", label = "Freshwater", x = -31, y = -29, size = 2, colour = "black")+ 
  #annotate("text", label = "Wild boar", x = -30, y = -27, size = 2, colour = "black")+ 
  annotate("text", label = "Ruminant", x = -29, y = -35, size = 3, colour = "black")+ 
  #annotate("text", label = "C3 nuts", x = -35, y = -33, size = 2, colour = "black")+
  annotate("text", label = "Rice", x = -34, y = -34, size = 3, colour = "black")+   coord_fixed(ratio = 1)+
  theme(axis.ticks = element_blank())+
  theme_bw(base_size = 10)+
  theme(strip.text.x = element_blank())+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
theme(legend.position="none")
p
```

#Create labels for individual plots
```{r}
# A data frame with labels for each facet
p_labels <- data.frame(Country= c('Korea', 'Japan'), label = c("Bronze Age Korea", "Jomon-Yayoi Japan"))
```

Plot millet biomarkers by 'culture'

```{r}
points <- c("Mumun" = 21, "Jomon" = 23, "Yayoi" = 21)
p1 <-p +
    geom_point(data=subset(Data, Millet=='n/a'), aes(y=d13C18, x=d13C16, shape=Phase_2), size=2, fill= "lightgrey", colour ="black")+
  geom_point(data=subset(Data, Millet=='no'), aes(y=d13C18,x=d13C16, shape= Phase_2), size=2, fill= "white",colour ="black")+
  geom_point (data= subset(Data, Millet=="yes"), aes(y=d13C18, x=d13C16, shape=Phase_2), fill= "#E69F00", size=2, colour="black")+
  scale_shape_manual(values=points)
p1 <- p1 + facet_wrap (~factor(Country, levels=c('Korea', 'Japan')))
p1 +  geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)
p1 + geom_text(data=subset(Data, Type=='Foodcrust'), aes(y=d13C18,x=d13C16, label = "*"),vjust = 0.2, size=4) 
```



Plot aquatic biomarkers by 'culture'. 
```{r}
p2 <-p +
  geom_point(data=subset(Data, Aquatic=='n/a'), aes(y=d13C18,x=d13C16, shape=Phase_2),  size=2, fill= "lightgrey", colour ="black")+ 
    geom_point(data=subset(Data, Aquatic=='no'), aes(y=d13C18,x=d13C16, shape=Phase_2), size=2,fill= "white", colour ="black")+ 
  geom_point (data= subset(Data, Aquatic=="yes"), aes(y=d13C18, x=d13C16, shape=Phase_2), fill= "#56B4E9", size=2, colour="black")+
  scale_shape_manual(values=points)
p2 <- p2 + facet_wrap(~factor(Country, levels=c('Korea', 'Japan')))
p2 +   geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)
p2 + geom_text(data=subset(Data, Type=='Foodcrust'), aes(y=d13C18,x=d13C16, label = "*"),vjust = 0.2, size=4) 

```

# Create Fig 2. 

```{r}
library(ggpubr)
Fig2 <- ggarrange(p1, p2,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2, widths = c(1, 2))
ggsave("Fig2.png", width = 17.8, height = 17.8, dpi = 300, units = "cm", device='png')
Fig2
```





Credible limits for rice lipids. The shaded color scale shows the maximum credible contribution (95% confidence) of each product (expressed as % fatty acid to total fatty acid). 

```{r}
Millet_pots <-Data_pots %>% filter (Category == "Millet" )
p3 <- p +
      geom_point(data = Millet_pots, aes(y=d13C18, x=d13C16,fill=p95, shape=Phase_2), size=2)+
    scale_fill_gradient (low="blue", high ="orange", limits = c(0,1), labels = label_percent(), name = "maximum\ncontribution(%)")+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
  p3 <- p3 +
      scale_shape_manual(values=points)+
  geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)+
    guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(3, "cm"))
  p3
  
```
```{r}
Marine_pots <-Data_pots %>% filter (Category == "Marine" )
p4 <- p +
      geom_point(data = Marine_pots, aes(y=d13C18, x=d13C16,fill=p95, shape=Phase_2), size=2)+
    scale_fill_gradient (low="blue", high ="orange", limits = c(0,1), labels = label_percent(), name = "maximum\ncontribution(%)")+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
  p4 <- p4 +
      scale_shape_manual(values=points)+
  geom_text(x = -30, y = -16, aes(label = label), data = p_labels, size=3)+
    guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(3, "cm"))
  p4
```

```{r}
Rice_pots <-Data_pots %>% filter (Category == "Rice" )
p5 <- p +
      geom_point(data = Rice_pots, aes(y=d13C18, x=d13C16,fill=p95, shape=Phase_2), size=2,show.legend =TRUE)+
    scale_fill_gradient (low="blue", high ="orange", limits = c(0,1), labels = label_percent(), name = "maximum\ncontribution(%)")+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
   p5 <-p5 +
  scale_shape_manual(values=points)+
  geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)+
   guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(2, "cm"))
   p5
```

```{r}
Rum_pots <-Data_pots %>% filter (Category == "Wild ruminant" )
p6 <- p +
      geom_point(data = Rum_pots, aes(y=d13C18, x=d13C16,fill=p95, shape=Phase_2), size=2)+
    scale_fill_gradient (low="blue", high ="orange", limits = c(0,1))+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
 p6 <- p6 +
     scale_shape_manual(values=points)+
  geom_text(x = -30, y = -16, aes(label = label), data = p_labels, size=3)+
    guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(2, "cm"))
 p6
```



#6. Create Fig 3. Mixing model

```{r}
library(ggpubr)
Fig3 <- ggarrange(p3,p4,p5, labels = c("A","B","C"), ncol = 1, nrow = 3, common.legend = TRUE, legend="bottom")
Fig3 
ggsave("Fig3.png", width = 17, height = 27, dpi = 300, units = "cm", device='png')
```







```{r}
Rice_pots <-Data_pots %>% filter (Category == "Rice" )
p4 <- p +
      geom_point(data = Rice_pots, aes(y=d13C18, x=d13C16,fill=p5, shape=Phase_2), size=2,show.legend =TRUE)+
    scale_fill_gradient (low="blue", high ="orange", limits = c(0,1), labels = label_percent(), name = "minimum rice\ncontribution(%)")+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
  p4 +
  scale_shape_manual(values=points)+
  geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)+
    guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(2, "cm"))
```
#Plot data against the E/H ratio 

```{r}

p5 <- p +
      geom_point(data = Data, aes(y=d13C18, x=d13C16, fill=EH, shape=Phase_2), size=2,show.legend =TRUE)+
    scale_fill_gradient (low="blue", high ="orange",  limits = c(2,8), name = "APAA E/H ratio")+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
  p5 +
  scale_shape_manual(values=points)+
  geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)+
    guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(2, "cm"))
```
#Plot data against the %SRR

```{r}

p5 <- p +
      geom_point(data = Data, aes(y=d13C18, x=d13C16, fill=SRR, shape=Phase_2), size=2,show.legend =TRUE)+
    scale_fill_gradient (low="blue", high ="orange",  limits = c(50,100), name = "% Phytanic acid SRR ratio")+
  #labs(fill = "Rice") +
  #theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_wrap (~factor(Country, levels=c('Korea', 'Japan'))) 
  p5 +
  scale_shape_manual(values=points)+
  geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)+
    guides(shape = "none") +
    theme(legend.position="right", legend.title = element_text(size = 8), legend.spacing.y = unit(2, "cm"))
```






#7 Plot sites spatially

```{r}
#Plot site location on map. Install packages and draw base map
#install.packages("sf", type = "source", configure.args = "--with-proj-lib=$(brew --prefix)/lib/")
library(devtools)
library(sf)
library(elevatr)
library(marmap)
#devtools::install_github("ropensci/rnaturalearthhires",auth_token = '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', force = TRUE)
library("rnaturalearth")
library(ggplot2)
library(raster)
library(gridExtra)
library(progress)
```


```{r}
#Draw basemap
win  <- ne_countries(continent = 'asia',scale=10,returnclass='sf')
japan <- ne_states(country = "japan",returnclass='sf')
elevation <- get_elev_raster(locations = win, z = 5,clip = "locations",src='aws')
dem <- as.data.frame(elevation, xy = T) |> na.omit()
colnames(dem)[3] <- "elevation"
dem$elevation[which(dem$elevation < 0)] = 0
```
Define datasets for plotting
```{r}
Map <-read_sheet('https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0', sheet="Map_coords") |> as.data.frame()
Pots <- Map %>% filter (Type == "Pot")
Bones <- Map %>% filter (Type == "Bone")

pots.sf <- st_as_sf(Pots, coords=c('Long','Lat'),crs=4326)
bones.sf  <- st_as_sf(Bones, coords=c('Long','Lat'),crs=4326)

```



```{r}
library("ggrepel")
library("ggspatial")

plot <- ggplot() +
	geom_tile(data=dem,aes(x=x,y=y,fill=elevation)) +
	scale_fill_etopo() +
	geom_sf(data=pots.sf, fill="#E69F00", shape=21, colour = "black", size=2) +
  geom_sf(data=bones.sf, fill="#56B4E9", shape=21, colour = "black", size=2)+
  geom_text_repel(data = Map, aes(x = Long, y = Lat, label = Num), nudge_x =  Map$Nudge_x, nudge_y = Map$Nudge_y, size=3) +
	xlim(124,135) +
	ylim(30,40) +
	xlab('Longitude') +
	ylab('Latitude') +
  annotation_scale()+
	theme(axis.ticks = element_blank())+
  theme_bw(base_size = 10)+
  theme(strip.text.x = element_blank())+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
theme(legend.position="none")
plot
```






Plot millet locations spatially. 

```{r}
Millet_locations  <-read_sheet('https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0', sheet="Botanical_AMS") |> as.data.frame()
Millet_locations <- Millet_locations %>% filter (Period == "Mumun" | Period == "Yayoi"| Period == "Jomon")
Panicum_locations <- Millet_locations %>% filter (PM == "Pa" )
Setaria_locations <- Millet_locations %>% filter(SI == "Se")
Panicum.sf  <- st_as_sf(Panicum_locations,coords=c('Longitude','Latitude'),crs=4326)
Setaria.sf  <- st_as_sf(Setaria_locations,coords=c('Longitude','Latitude'),crs=4326)
Millet.sf  <- st_as_sf(Millet_locations,coords=c('Longitude','Latitude'),crs=4326)

```

```{r}
plot <- ggplot() +
	geom_tile(data=dem,aes(x=x,y=y,fill=elevation)) +
	scale_fill_etopo() +
	geom_sf(data=Panicum.sf, size=2,col='black',pch=20) +
  geom_sf(data=Millet.sf, size=2,pch=20, fill=Millet) +
	xlim(125,135) +
	ylim(30,40) +
	xlab('Longitude') +
	ylab('Latitude') +
	theme(legend.position = "none")
plot
```

#8 Plot human isotope data


```{r}
Hum <-read_sheet('https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0', sheet="hum_isotopes") |> as.data.frame()
```
```{r}
Data_hum <-read_sheet('https://docs.google.com/spreadsheets/d/1vdw8Osu5heuTHNIOzy9qM8NY-qKUZDM12JsbqVJYS20/edit#gid=0', sheet="hum_isotopes") |> as.data.frame()

```
```{r}
p <-ggplot()+
  labs(y=expression(delta^{15}*N[collagen]*"(\u2030)"), x=expression(delta^{13}*C[collagen]*"(\u2030)"))+
  scale_x_continuous(limits=c(-25,-10))+
  scale_y_continuous(limits=c(0,15))+
  geom_point(data=Data_hum, aes(y=d15N_coll, x=d13C_coll,  fill= Period_2), size=3, pch= 21,  colour="black")+
  theme(axis.ticks = element_blank())+
  theme_bw(base_size = 10)+
  theme(strip.text.x = element_blank())+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
theme(legend.position="none")+
  
p
```


```{r}
p

```
p