---
title: "KyushuKorea_organicresidues"
author: "OEC"
date: "27/02/2024"
output: html_document
editor_options: 
  chunk_output_type: inline
---


#1. Preservation and summary statisitcs


#Read files 
```{r}
library(here)
Data  <-  read.csv(here('data','S1.csv'))
```

#1a. Preservation by site
```{r}
library(dplyr)
#Grouped by site
Conc_site <- Data %>% dplyr::filter(Sample_type == 'ceramic')%>%
  group_by(Site_code) %>% 
  summarise(n = n(), mean.conc = mean(Yield_AE),
            sd.conc = sd(Yield_AE), n_lipid = sum(Yield_AE >=5))%>%
  mutate("%interp"= n_lipid/n*100)
Conc_site <- Conc_site %>% arrange(factor(Site_code, levels = c('OBD', 'RMM', 'SHC', 'SAS','NBK', 'MSR', 'DPR', 'SDR', 'SWR', 'JKR', 'KCR', 'MJR')))
write.csv (Conc_site,"Site_Preservation.csv")#create csv output 
```
  
#1b. Preservation by region

```{r}
Region <-Data %>%filter(Sample_type == 'ceramic') %>% 
 group_by(Country) %>% 
  summarise(n = n(), median.conc = median(Yield_AE),
           mean.conc = mean(Yield_AE),
           sd.conc = sd(Yield_AE), n_lipid = sum(Yield_AE >=5)) %>%
  mutate("%interp"= n_lipid/n*100)
 write.csv (Region,"Region_preservation.csv")#create csv output 
```
  
#1c.Biomarker identification by site by interpretable lipid

```{r}
Inter_lipid <- Data %>%filter (Yield_AE >=5)
Biomarker <-Inter_lipid %>% group_by(Site_code) %>% 
  summarise (n = n(), n_millet = sum (Miliacin == "yes"), n_aquatic = sum (Aquatic == "yes")) %>%
  mutate("%millet"= n_millet/n*100, "%aqatiuc"= n_aquatic/n*100)
Biomarker <- Biomarker %>% arrange(factor(Site_code, levels = c('OBD', 'RMM', 'SHC', 'SAS', 'NBK', 'MSR', 'DPR', 'SDR', 'SWR', 'JKR', 'KCR', 'MJR')))
write.csv (Biomarker,"Biomarker_preservation.csv")#create csv output 
```


#2. Load and summarise isotope and isotopic reference data

#2a.Isotpic Reference data

Obtain data for isotope references and concentations
```{r}
#Otain data for reference fats
library(dplyr)
Iso_ref <- read.csv(here('data','S2.csv')) %>% filter(Category %in% c('Rice', 'Millet', 'Wild ruminant', 'Marine'))

Conc_ref <- read.csv(here('data','S3.csv')) %>% filter(Category %in% c('Rice', 'Millet', 'Wild ruminant', 'Marine'))
```

Find means and ranges of isotope and concentrations values
```{r}
Iso_sum <- Iso_ref %>% group_by(Category) %>% summarise (Meand13C16 = mean(d13C16), SDd13C16 = sd(d13C16), Meand13C18 = mean(d13C18), SDd13C18 = sd(d13C18), n = n())

Conc_sum <- Conc_ref %>% group_by(Category) %>% summarise (MeanC16 = mean(Ab_Percent_C16), SDC16 = sd(Ab_Percent_C16), MeanC18 = mean(Ab_Percent_C18), SDC18 = sd(Ab_Percent_C18), n_conc = n())

```

#2b Randomly sample  concentration values for each isotope value based on the mean and sd of the concentration 

```{r}
library(purrr)
library(tidyr)
Conc <- inner_join(Iso_sum, Conc_sum, 
           by = c("Category"))
# Generate the random values using purrr::map and bind them into a single column
Conc <- Conc %>%
  # Use purrr::pmap to generate random numbers based on each row's parameters
  mutate(
    Concd13C16 = pmap(list(MeanC16, SDC16, n), function(mean, sd, n) {
      rnorm(n, mean = mean, sd = sd)
    }),
    Concd13C18 = pmap(list(MeanC18, SDC18, n), function(mean, sd, n) {
      rnorm(n, mean = mean, sd = sd)
    })
  ) %>%
  unnest(cols =c(Concd13C16, Concd13C18))
  # Unnest the list column to create one long column of generated values
```

#3. Mixing model

#3a Load the consumer data
```{r}
#remotes::install_github("brianstock/MixSIAR")
library(MixSIAR)

here('data','S1.csv')

Pots <- Data %>% filter(d13C16 < 0) %>% dplyr::select(Sample_ID, Site_name, Region, Phase_2, d13C16, d13C18)
write.csv(Pots,file=here::here("organicresidues", "pots.csv"))
mix <- load_mix_data (here("organicresidues","pots.csv"), 
                     iso_names=c("d13C16", "d13C18"), 
                     factors="Sample_ID",
                     fac_random=FALSE, 
                     fac_nested=FALSE, 
                     cont_effects=NULL)
```


#3b. Create a source data file

```{r}
#Select categories for model and index
Iso_ref1 <- Iso_ref %>% filter(Category == "Rice" | Category == "Millet" | Category == "Wild ruminant" | Category == "Marine") %>% mutate ("Sources"= Category) %>%dplyr::select(Sources, d13C16, d13C18) 
Iso_ref1 <- Iso_ref1 %>%
  arrange(Sources) %>%
  mutate(Observation = row_number()) %>%
  ungroup()
#Repeat for concentration
Conc_ref1 <- Conc %>% filter(Category == "Rice" | Category == "Millet" | Category == "Wild ruminant" | Category == "Marine") %>% mutate ("Sources"= Category) %>%dplyr::select(Sources, Concd13C16, Concd13C18) 
Conc_ref1 <- Conc_ref1 %>%
  arrange(Sources) %>%
  mutate(Observation = row_number()) %>%
  ungroup()
#Join. the data togeter
Source <- Source <- inner_join(Conc_ref1, Iso_ref1, by = c("Observation"))
Source <- Source %>% dplyr::select(-c(Sources.y, Observation))#drop unwanted columns
colnames(Source)[1] <- "Sources"

```
#Write and load the source data

```{r}
write.csv (Source, file=here::here("organicresidues", "source.csv"), row.names=FALSE)
source <- load_source_data (here::here("organicresidues", "source.csv"), 
                           source_factors=NULL,
                             conc_dep=TRUE, 
                           data_type="raw", 
                           mix)
```

#3c Load the TDFs if relevant or nill return
```{r}
#create a dataframe with nill values for each source
Discr <- data.frame(
  Sources = c("Marine", "Millet", "Rice", "Wild ruminant"),
  Meand13C16 = c(0,0,0,0),	
  SDd13C16	= c(0,0,0,0),
  Meand13C18	=c (0,0,0,0),
  SDd13C18 = c(0,0,0,0))
write.csv (Discr, file=here::here("organicresidues", "Discr.csv"), row.names=FALSE)

# Load the discrimination/TDF data
discr <- load_discr_data(here::here("organicresidues", "Discr.csv"), mix)
```

#3d Create an isospace plot

```{r, eval=FALSE}
# Make an isospace plot
library (ggplot2)
plot <- plot_data(filename="isospace_plot", plot_save_pdf=F, plot_save_png=F,mix,source, discr, return_obj = TRUE)
plot  + theme(panel.border = element_blank(), panel.background = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
axis.line = element_line(colour = "black"),
axis.title=element_text(size=12), axis.text=element_text(size=12),
legend.position="none")
```
```{r}
# Calculate the convex hull area, standardized by source variance
calc_area(source=source,mix=mix,discr=discr)
```
Specify priors
```{r, eval=FALSE}
# default "UNINFORMATIVE" / GENERALIST prior (alpha = 1)
#plot_prior(alpha.prior=1,source)
```

#3e. Create the JAGS file and run the model. Choose process errors only although this needs to be considered carefully depending on the model. https://link.springer.com/article/10.1007/s10816-020-09492-5

```{r, eval=FALSE}
# Write the JAGS model file
model_filename <- "MixSIAR_model.txt"   # Name of the JAGS model file
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix, source)
```

```{r, eval=FALSE}
#run <- list(chainLength=200000, burn=150000, thin=50, chains=3, calcDIC=TRUE)
```

```{r, eval=FALSE}
jags.1 <- run_model(run="test", mix, source, discr, model_filename)
```
```{r, eval=FALSE}
jags.1 <- run_model(run="long", mix, source, discr, model_filename)
```

#3f. Obtain the model outputs
```{r, eval=FALSE}
output_options <- list(summary_save = TRUE,
                       summary_name = "summary_final",
                       sup_post = TRUE,
                       plot_post_save_pdf = FALSE,
                       plot_post_name = "posterior_density",
                       sup_pairs = TRUE,
                       plot_pairs_save_pdf = FALSE,
                       plot_pairs_name = "pairs_plot",
                       sup_xy = TRUE,
                       plot_xy_save_pdf = FALSE,
                       plot_xy_name = "xy_plot",
                       gelman = TRUE,
                       heidel = TRUE,
                       geweke = TRUE,
                       diag_save = TRUE,
                       diag_name = "diagnostics",
                       indiv_effect = FALSE,
                       plot_post_save_png = FALSE,
                       plot_pairs_save_png = FALSE,
                       plot_xy_save_png = FALSE,
                       diag_save_ggmcmc = FALSE,
                       return_obj=TRUE)
```

Get posterior chains into one tidy data frame
```{r}
# Get posterior chains into one tidy data frame
library(R2jags)

summary(jags.1)
attach.jags(jags.1)
```
Run diagnostics and create a data frame with some summary stats. 

```{r}
diag <- output_diagnostics(jags.1, mix, source, output_options)
head(diag$gelman)
df.stats <- output_stats(jags.1, mix, source, output_options)
rownames(df.stats)
df.stats <- as.data.frame(df.stats)
diag
```

Add rownames to outputs
```{r}
df.stats <- cbind(Sample_category = rownames(df.stats), df.stats)
rownames(df.stats) <- 1:nrow(df.stats)
df.stats
```

Fix column names for visualization
```{r}
library(tidyr)
df.stats <- df.stats %>% separate_wider_delim (Sample_category, ".", names = c("number", "Sample_ID", "Category"))
write.csv(df.stats,(here('data','S6.csv')))#write csv file for supplementary 
colnames(df.stats)[11] <- "p95"
colnames(df.stats)[7] <- "p5"
```

Join with isotope data to create a dataframe for visualisation
```{r}
Iso_data <- Data %>% filter(d13C16 < 0) 
Data_pots <- inner_join(Iso_data, df.stats, 
           by = c("Sample_ID"))
```

Output posteriors 

```{r}
g.post <- output_posteriors(jags.1, mix, source, output_options)
```


#4. Create Figure 2

#4a. Create a base plot

```{r}
library("ggplot2")
p <-ggplot()+
  labs(y=expression(delta^{13}*C[18:0]*"(\u2030)"), x=expression(delta^{13}*C[16:0]*"(\u2030)"))+
  scale_x_continuous(limits=c(-37,-15))+
  scale_y_continuous(limits=c(-37,-15))+
     stat_ellipse(geom= "polygon", data=Iso_ref1, alpha=0.2, colour="lightgrey",size=0.1,
               aes (x = d13C16, y = d13C18, group = Sources),
              level = 0.68)+
annotate("text", label = "Millet", x = -18, y = -16, size = 3, colour = "black")+ 
  annotate("text", label = "Marine", x = -24, y = -21, size = 3, colour = "black")+ 
  annotate("text", label = "Ruminant", x = -29, y = -35, size = 3, colour = "black")+ 
  annotate("text", label = "Rice", x = -35.5, y = -34, size = 3, colour = "black")+   
  coord_fixed(ratio = 1)+
  theme(axis.ticks = element_blank())+
  theme_bw(base_size = 10)+
  theme(strip.text.x = element_blank())+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
theme(legend.position="none")
p
```

#4b Create labels for individual plots
```{r}
# A data frame with labels for each facet
p_labels <- data.frame(Country= c('Korea', 'Japan'), label = c("Bronze Age Korea", "Jomon-Yayoi Japan"))
points <- c("Mumun" = 21, "Jomon" = 23, "Yayoi" = 21)
```


#4c Plot millet biomarkers by 'culture'

```{r}
p1 <-p +
    geom_point(data=subset(Data, Millet=='n/a'), aes(y=d13C18, x=d13C16, shape=Phase_2), size=1.5, fill= "lightgrey", colour ="black")+
  geom_point(data=subset(Data, Miliacin=='no'), aes(y=d13C18,x=d13C16, shape= Phase_2), size=1.5, fill= "white",colour ="black")+
  geom_point (data= subset(Data, Miliacin=="yes"), aes(y=d13C18, x=d13C16, shape=Phase_2), fill= "#E69F00", size=1.5, colour="black")+
  scale_shape_manual(values=points)+
facet_wrap (~factor(Country, levels=c('Korea', 'Japan')))+
geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)
#geom_text(data=subset(Data, Type=='Foodcrust'), aes(y=d13C18,x=d13C16, label = "*"),vjust = 0.2, size=4) 
p1
```


#5c Plot aquatic biomarkers by 'culture'. 
```{r}
p2 <-p +
  geom_point(data=subset(Data, Aquatic=='n/a'), aes(y=d13C18,x=d13C16, shape=Phase_2),  size=1.5, fill= "lightgrey", colour ="black")+ 
    geom_point(data=subset(Data, Aquatic=='no'), aes(y=d13C18,x=d13C16, shape=Phase_2), size=1.5,fill= "white", colour ="black")+ 
  geom_point (data= subset(Data, Aquatic=="yes"), aes(y=d13C18, x=d13C16, shape=Phase_2), fill= "#56B4E9", size=1.5, colour="black")+
  scale_shape_manual(values=points)
p2 <- p2 + facet_wrap(~factor(Country, levels=c('Korea', 'Japan')))
p2 +   geom_text(x = -32, y = -16, aes(label = label), data = p_labels, size=3)
#geom_text(data=subset(Data, Type=='Foodcrust'), aes(y=d13C18,x=d13C16, label = "*"),vjust = 0.2, size=4) 

```

#5d. Save biomarker plot. 

```{r}
library(ggpubr)
Fig2a <- ggarrange(p1, p2,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2, widths = c(1, 2))
ggsave("Fig2a.svg", width = 11.4, height = 11.5, dpi = 300, units = "cm", device='svg')
Fig2a
```

#5e. Visalise the outputs of the mixing model

```{r}
library("scales")
Data_pots_dropWR <-Data_pots %>% filter (Category != 'Wild ruminant')#drop Wild Ruminant
Data_pots_dropWR$Country =  factor(Data_pots_dropWR$Country, levels=c('Korea', 'Japan'))#order country variable
Data_pots_dropWR$Category =  factor(Data_pots_dropWR$Category, levels=c('Millet', 'Marine', 'Rice'))#order category variable
Fig2b <- p +
      geom_point(data = Data_pots_dropWR, aes(y=d13C18, x=d13C16,fill=p95, shape=Phase_2), width=1, size=1)+
    scale_fill_gradient (low="blue", high ="orange", limits = c(0,1), labels =   label_percent(), name = "maximum\ncontribution(%)")+
  theme(legend.position = 'bottom', legend.direction = "horizontal")+
  facet_grid (Category ~ Country) +
    scale_shape_manual(values=points)+
    guides(shape = "none")
ggsave("Fig2b.svg", width =11.4, height = 22, dpi = 300, units = "cm", device='svg')
Fig2b
```

